\tool: Programmable, Efficient, and Scalable Secure Two-Party Computation

### 2017-10-29 OaklandSubmission/ld.tex
temp

\newcommand{\kw}[1]{{\lstinline[basicstyle=\small\color{blue}]{#1}}}
%\newcommand{\kw}[1]{{\ensuremath{\mathtt{#1}}}}
\newcommand{\ftext}[1]{\text{\small{#1}}}
\newcommand{\cond}[3]{\ensuremath{{{#1}\:?\:{#2}\::{#3}}}}
\newcommand{\for}[4]{\ensuremath{\kw{for}\:{#1}\:\kw{in}\:[{#2}, {#3}]\:\kw{do}\:{#4}}}
\newcommand{\ite}[3]{\ensuremath{\kw{if}({#1}, {#2}, {#3})}}
\newcommand{\loops}[3]{\ensuremath{\kw{while}\:{#1} \leq {#2}\:\kw{do}\:{#3}}}

@@

In this section we prove correctness and security of our compiler \tool.
We formalize our source, intermediate and circuit languages. 
An example of a program in our source language is provided in \figureref{ex-sml}.
%Our source runtime semantics is a model of the ideal, trusted third-party semantics, and
%generates observations corresponding to the values revealed to the
%parties.
Our source runtime semantics describe the standard execution semantics of the source programs and
generates observations corresponding to the values revealed to the
parties.
%
We present the compilation rules that type check a program in the source language and generates a program in the intermediate language. This intermediate language models programs such as \figureref{ex-aby} generated by our compiler implementation.
%
We evaluate the program in our intermediate language  to a circuit by ``computing away'' the public parts and the arrays. Crucially, this step does not have access to the secret inputs--those are processed by the distributed circuit semantics that model the \mpc back-end.
%
Here, the generated circuit is evaluated in the distributed setting, where the parties run an interactive protocol. This step, like  the source semantics, emits observations corresponding to the values revealed to the parties. 
%Finally, we formalize the distributed circuit semantics (of the cryptographic back-end) that computes the generated circuit in the distributed setting, and like  the source semantics, emits observationscorresponding to the values revealed to the parties.

To prove the correctness of \tool, we prove that  the observations in source semantics and distributed circuit semantics are identical (see \theoremref{correctness}). 
We combine this correctness and the security of the \mpc back-end to argue security of the implementations generated by \tool (see \theoremref{security}).
%For the security of our compiler, we rely on the correctness guarantee and  reduce the security to the security of the underlying cryptographic protocol used to compute the generated circuit (see \theoremref{security}).
We present only selected parts of our formalization for space
reasons. Full definitions and proofs can be found \cite{anonymized}.

@@

\begin{figure}
  \small
  \[
  \begin{array}{rrcl}
    \ftext{Base type} & \sigma &::=& \kw{uint} \mid \kw{bool}\\
    \ftext{Type} & \psi &::=& \sigma \mid \sigma[n]\\
    \ftext{Constant} & c &::=& n \mid \top \mid \bot\\
    \ftext{Expression} & e &::=& c \mid x \mid e_{1} \mulop e_{2} \mid e_{1} > e_{2} \mid \cond{e_{1}}{e_{2}}{e_{3}}\\
    & &\mid& [\overline{e_{i}}]_{n} \mid x[e] \mid \kw{in}_{j}\\
    \ftext{Statement} & s &::=& \psi\:x = e \mid x := e \mid \for{x}{n_{1}}{n_{2}}{s}\\
    & & \mid& x[e_{1}] := e_{2} \mid \ite{e}{s_{1}}{s_{2}} \mid \kw{out}\:e \mid s_{1}; s_{2}\\
    & & \mid& \loops{x}{n}{s}
  \end{array}
  \]
\caption{Source language}
\label{fig:srclang}
\end{figure}

@@

\subsubsection*{Source language} Our source language
 is a simple imperative language. \divya{We present the core of our language in \figureref{srclang} but our implementation is more expressive (see \sectionref{impl}).} Types
$\psi$ consist of the base types $\sigma$, and arrays
of base types $\sigma[n]$, where $n$ is the array length. Though we
model only one dimensional arrays, our implementation supports higher
dimensional arrays as well. Expressions $e$ in the language include the
integer constants $n$, \kw{bool} constants $\top$ and $\bot$,
variables $x$, binary operations $e_{1} \mulop e_{2}$ and $e_{1} > e_{2}$,
conditionals $\cond{e}{e_{1}}{e_{2}}$, array literals 
$[\overline{e_{i}}]_{n}$\footnote{We write $\overline{e}$ (and
  similarly for other symbols) to denote a sequence of expressions.
The length of the sequence is usually clear from the context.}, and
array reads $x[e]$. The expression $\kw{in}_{j}$ denotes input from
party $j$. The statements $s$ in the language comprise of variable
declarations and assignments ($\psi\:x = e$ and $x := e$ resp.),
\kw{for} loops, array writes ($x[e_{1}] := e_{2}$), \kw{if}
statements, and sequence of statements ($s_{1}; s_{2}$). The statement
$\kw{out}\:e$ denotes revealing the value of $e$ to the
parties. The \kw{while} statement is an internal syntax that is not
exposed to the programmer. \\

\noindent\textbf{Source Semantics.} The runtime semantics for the source language is shown in
\figureref{srcsem}. \TODO{Add something about claer text stand alone exec.} Values $v$, runtime environments $\rho$, and
observations $O$ are defined as follows:

@@

The judgment $\rho \vdash e \Downarrow v$ denotes the big-step
evaluation of an expression~$e$ to a value~$v$ under the runtime
environment~$\rho$. Rule ({\sc{E-Var}}) looks up the value of $x$ in
the environment. Rule ({\sc{E-Mult}}) inductively evaluates $e_{1}$ and
$e_{2}$, and evaluates to their product. Rule ({\sc{E-Read}})
evaluates an array read operation. It first evaluates $x$ to an array
value $[\overline{c_{i}}]_{n_{1}}$, and $e$ to a \kw{uint} value
$n$. It then returns $c_{n}$, the $n$-th index value in the array,
provided $n < n_{1}$, the length of the array. Rule ({\sc{E-Inp}})
evaluates to some constant $c$ denoting party $j$'s input. We model
the inputs to be base constants, an array input can be written in the
language as $[\kw{in}_{j}]_{n}$, which can then evaluate using the
rule ({\sc{E-Arr}}). The remaining rules are straightforward, and are
elided for space reasons.

@@

\begin{figure}
  \small
  \[
  \begin{array}{rrcl}
    \ftext{Secret label} & m &::=& \mathcal{A} \mid \mathcal{B}\\
    \ftext{Label} & \ell &::=& \mathcal{P} \mid m\\
    \ftext{Type} & \tau &::=& \sigma^{\ell} \mid \sigma^{\ell}[n]\\
    \ftext{Expression} & \widetilde{e} &::=& c \mid x \mid \widetilde{e_{1}} \mulop_{\ell} \widetilde{e_{2}} \mid \widetilde{e_{1}} >_{\ell} \widetilde{e_{2}} \\
    & & \mid & \lcond{\ell}{{\widetilde{e}}}{{\widetilde{e_{1}}}}{{\widetilde{e_{2}}}} \mid x[\widetilde{e}] \mid [\overline{\widetilde{e_{i}}}]_{n} \mid \kw{in}^{m}_{j} \mid \widetilde{e} \rhd m\\
    \ftext{Statement} & \widetilde{s} &::=& \tau\:x = \widetilde{e} \mid x := \widetilde{e} \mid \dots \mid \widetilde{s_{1}}; \widetilde{s_{2}} \mid \dots\\
  \end{array}
  \]
\caption{Intermediate Language}
\label{fig:interlang}
\end{figure}

\subsubsection*{Intermediate language} \figureref{interlang} shows the
intermediate language of our compiler. The syntax follows that of the source
language, except that the types and operators are \emph{labeled}. Below
we mainly focus on the bits that are different from the source
language.

Labels $\ell$ consist of secret labels $\mathcal{A}$ and
$\mathcal{B}$, denoting the arithmetic and boolean shared secrets
resp., and the public label $\mathcal{P}$. Types $\tau$ are then
labeled base types $\sigma^{\ell}$ and arrays of labeled base types
$\sigma^{\ell}[n]$.

Most of the expression forms $\widetilde{e}$ are same as $e$, except
that the binary operators, and the conditional forms
are annotated with label $\ell$, denoting how the operators should be
evaluated ($\mathcal{P}$ for in-clear, and $\mathcal{A}$ and
$\mathcal{B}$ for using arithmetic or boolean circuits resp.). The
form $\widetilde{e} \rhd m$ denotes coercing the value of $\widetilde{e}$ to be
$m$-secret shared. The statements $\widetilde{s}$ are analogous to
$s$.

@@

	\inferrule*[lab={\footnotesize{T-Read}}]
               {
                 \Gamma \vdash x : \sigma^{\ell}[n] \leadsto x\\\\
                 \Gamma \vdash e : \kw{uint}^{\mathcal{P}} \leadsto \widetilde{e}\\\\
                 \Gamma \models e < n
               }
               {
                 \Gamma \vdash x[e] : \sigma^{\ell} \leadsto x[\widetilde{e}]
               }

@@

     \inferrule*[lab={\footnotesize{T-Write}}]
               {
                 \Gamma \vdash x : \sigma^{\ell}[n] \leadsto x\\\\
                 \Gamma \vdash e_{1} : \kw{uint}^{\mathcal{P}} \leadsto \widetilde{e_{1}}\\\\
                 \Gamma \vdash e_{2} : \sigma^{\ell} \leadsto \widetilde{e_{2}}\\\\
                 \Gamma \models e_{1} < n
               }
               {
                 \Gamma \vdash x[e_{1}] := e_{2} \leadsto x[\widetilde{e_{1}}] := \widetilde{e_{2}} \mid \Gamma
               }

@@

\subsubsection*{Source to Intermediate Compilation judgments} We provide the compilation
rules in \figureref{compile}. \divya{PL jargon?} We present the rules in a
declarative style, where the rules are non-syntax
directed, and the labels $\ell$ are chosen
non-deterministically. \sectionref{impl} describes the label  inference
scheme in our implementation.

Judgment $\Gamma \vdash e : \tau \leadsto \widetilde{e}$, where
$\Gamma$ maps variables $x$ to types $\tau$, says that
under $\Gamma$, $e$ (in source language) compiles to $\widetilde{e}$ (in intermediate language) with type $\tau$.

The rule ({\sc{T-Cons}}) assigns the label $\mathcal{P}$ to the
constants, as the constants are always public. Rule ({\sc{T-Mult}})
compiles a multiplication to either a public multiplication  ($\mulop_{\mathcal{P}}$), or a
secret arithmetic multiplication ($\mulop_{\mathcal{A}}$). As
our compiler is cryptographic cost aware, it never compiles the
multiplication to boolean multiplication $\mulop_{\mathcal{B}}$. In a similar manner,
rule ({\sc{T-Gt}}) compiles $e_{1} > e_{2}$ to either public comparison,
or secret boolean comparison. The rule for conditional
({\sc{T-Cond}}) has two cases: when the conditional expression $e$ is of
type $\kw{bool}^{\mathcal{P}}$, both the branches have a base type
$\sigma^{\ell'}$, for an arbitrary $\ell'$, and the conditional is
compiled to a public conditional, whereas when the conditional
expression has type $\kw{bool}^{\mathcal{B}}$, $\ell'$ is also
$\mathcal{B}$, and the conditional is compiled to a secret conditional
using a boolean circuit. Note that we restrict the type of the branches
to be of base type. 
Rule ({\sc{T-Read}}) type checks an array read. It checks
that the array index $e$ is public, and uses a static bounds checking
judgment $\Gamma \models e < n$ to prove that the array index is in
bounds. Section~\ref{sec:impl} discusses our implementation of bounds
checking. Rule ({\sc{T-Inp}}) picks a label $m$ for the input. Finally,
rule ({\sc{T-Sub}}) is the subsumption rule that coerces an expression of
type $\sigma^{\ell}$ to an expression of type $\sigma^{m}$ using the
coerce expression. It is important for security that the secrets cannot be coerced to public
values.

>>>>

%\newcommand{\kw}[1]{{\lstinline[basicstyle=\small\color{blue}]{#1}}}
\newcommand{\kw}[1]{{\ensuremath{\mathtt{#1}}}}
\newcommand{\ftext}[1]{\text{\small{#1}}}
\newcommand{\cond}[3]{\ensuremath{{{#1}\:?\:{#2}\::{#3}}}}
\newcommand{\for}[4]{\ensuremath{\kw{for}\:{#1}\:\kw{in}\:[{#2}, {#3}]\:\kw{do}\:{#4}}}
\newcommand{\ite}[3]{\ensuremath{\kw{if}({#1}, {#2}, {#3})}}
\newcommand{\loops}[3]{\ensuremath{\kw{while}\:{#1} \leq {#2}\:\kw{do}\:{#3}}}

@@

In this section we prove correctness and security of our \tool compiler.
We formalize our source, intermediate and circuit languages. 
An example of a program in our source language is provided in \figureref{ex-sml}.
%Our source runtime semantics is a model of the ideal, trusted third-party semantics, and
%generates observations corresponding to the values revealed to the
%parties.
Our source runtime semantics describe the ``trusted third party" execution semantics of the source programs and
generate observations corresponding to the values revealed to the
parties.
%
We present the compilation rules that type check a program in the source language and generate a program in the intermediate language. This intermediate language models programs such as \figureref{ex-aby} generated by our compiler implementation.
%
We evaluate the program in our intermediate language  to a circuit by ``evaluating away'' the public parts and the arrays. Crucially, this step does not have access to the secret inputs; those are processed by the distributed circuit semantics that model the \mpc back-end.
%
Evaluation in the distributed setting involves the parties running an interactive protocol. This step, like  the source semantics, emits observations corresponding to the values revealed to the parties. 
%Finally, we formalize the distributed circuit semantics (of the cryptographic back-end) that computes the generated circuit in the distributed setting, and like  the source semantics, emits observationscorresponding to the values revealed to the parties.

To prove the correctness of \tool, we prove that  the observations in source semantics and distributed circuit semantics are identical (see \theoremref{correctness}). 
We combine this correctness guarantee and the security of the \mpc back-end to argue security of the implementations generated by \tool (see \theoremref{security}).
%For the security of our compiler, we rely on the correctness guarantee and  reduce the security to the security of the underlying cryptographic protocol used to compute the generated circuit (see \theoremref{security}).
We present only selected parts of our formalization for space
reasons. Full definitions and proofs can be found in~\cite{anonymized}.

@@

\begin{figure}
  \small
  \[
  \begin{array}{rrcl}
    \ftext{Base type} & \sigma &::=& \kw{uint} \mid \kw{bool}\\
    \ftext{Type} & \psi &::=& \sigma \mid \sigma[n]\\
    \ftext{Constant} & c &::=& n \mid \top \mid \bot\\
    \ftext{Expression} & e &::=& c \mid x \mid e_{1} \mulop e_{2} \mid e_{1} > e_{2} \mid \cond{e_{1}}{e_{2}}{e_{3}}\\
    & &\mid& [\overline{e_{i}}]_{n} \mid x[e] \mid \kw{in}_{j}\\
    \ftext{Statement} & s &::=& \psi\:x = e \mid x := e \mid \for{x}{n_{1}}{n_{2}}{s}\\
    & & \mid& x[e_{1}] := e_{2} \mid \ite{e}{s_{1}}{s_{2}} \mid \kw{out}\:e \mid s_{1}; s_{2}\\
    & & \mid& \loops{x}{n}{s}
  \end{array}
  \]
\caption{Source language syntax}
\label{fig:srclang}
\end{figure}

@@

\subsubsection*{Source language} Our source language
 is a simple imperative language and we show a subset of the supported operators in \figureref{srclang}. Types
$\psi$ consist of the base types $\sigma$, and arrays
of base types $\sigma[n]$, where $n$ is the array length. Though we
model only one dimensional arrays, our implementation supports
 multi-dimensional arrays as well. Expressions $e$ in the language include the
 integer constants $n$, \kw{bool} constants $\top$ and $\bot$,
variables $x$, binary operations $e_{1} \mulop e_{2}$ and $e_{1} > e_{2}$,
conditionals $\cond{e}{e_{1}}{e_{2}}$, array literals 
$[\overline{e_{i}}]_{n}$\footnote{We write $\overline{e}$ (and
  similarly for other symbols) to denote a sequence of expressions.
The length of the sequence is usually clear from the context.}, and
array reads $x[e]$. The expression $\kw{in}_{j}$ denotes input from
party $j$. The statements $s$ in the language comprise of variable
declarations and assignments ($\psi\:x = e$ and $x := e$ resp.),
\kw{for} loops, array writes ($x[e_{1}] := e_{2}$), \kw{if}
statements, and sequence of statements ($s_{1}; s_{2}$). The statement
$\kw{out}\:e$ denotes revealing the value of $e$ to the
parties. The \kw{while} statement is an internal syntax that is not
exposed to the programmer. \\

\noindent\textbf{Source semantics.} The runtime semantics for the source language is shown in
\figureref{srcsem}. These semantics show how a ``trusted third party" computes the outputs when given the inputs of both the parties. 
 Values $v$, runtime environments $\rho$, and
observations $O$ are defined as follows:

@@

The judgment $\rho \vdash e \Downarrow v$ denotes the big-step
evaluation of an expression~$e$ to a value~$v$ under the runtime
environment~$\rho$. Rule ({\sc{E-Var}}) looks up the value of $x$ in
the environment. Rule ({\sc{E-Mult}}) inductively evaluates $e_{1}$ and
$e_{2}$, and evaluates to their product. Rule ({\sc{E-Read}})
evaluates an array read operation. It first evaluates $x$ to an array
value $[\overline{c_{i}}]_{n_{1}}$, and $e$ to a \kw{uint} value
$n$. It then returns $c_{n}$, the $n$-th index value in the array,
provided $n < n_{1}$, the length of the array. Rule ({\sc{E-Inp}})
evaluates to some constant $c$ denoting party $j$'s input. 
%We model the inputs to be base constants, a
An array input can be written in the
language as $[\kw{in}_{j}]_{n}$, which can then evaluate using the
rule ({\sc{E-Arr}}). The remaining rules are straightforward, and are
elided for space reasons.

@@

\begin{figure}
  \small
  \[
  \begin{array}{rrcl}
    \ftext{Secret label} & m &::=& \mathcal{A} \mid \mathcal{B}\\
    \ftext{Label} & \ell &::=& \mathcal{P} \mid m\\
    \ftext{Type} & \tau &::=& \sigma^{\ell} \mid \sigma^{\ell}[n]\\
    \ftext{Expression} & \widetilde{e} &::=& c \mid x \mid \widetilde{e_{1}} \mulop_{\ell} \widetilde{e_{2}} \mid \widetilde{e_{1}} >_{\ell} \widetilde{e_{2}} \\
    & & \mid & \lcond{\ell}{{\widetilde{e}}}{{\widetilde{e_{1}}}}{{\widetilde{e_{2}}}} \mid x[\widetilde{e}] \mid [\overline{\widetilde{e_{i}}}]_{n} \mid \kw{in}^{m}_{j} \mid \widetilde{e} \rhd m\\
    \ftext{Statement} & \widetilde{s} &::=& \tau\:x = \widetilde{e} \mid x := \widetilde{e} \mid \dots \mid \widetilde{s_{1}}; \widetilde{s_{2}} \mid \dots\\
  \end{array}
  \]
\caption{Intermediate language syntax}
\label{fig:interlang}
\end{figure}

\subsubsection*{Intermediate language} \figureref{interlang} shows the
intermediate language of our compiler. The syntax follows that of the source
language, except that the types and operators are \emph{labeled}. 
%Below we mainly focus on the bits that are different from the source
%language.
A label $\ell$ can be the  public label $\mathcal{P}$ or  one of the secret labels $\mathcal{A}$ or
$\mathcal{B}$, which denote arithmetic and boolean respectively.
Types $\tau$ are then
labeled base types $\sigma^{\ell}$ and arrays of labeled base types
$\sigma^{\ell}[n]$.

Most of the expression forms $\widetilde{e}$ are same as $e$, except
that the binary operators, and the conditional forms
are annotated with labels $\ell$, denoting how the operators should be
evaluated ($\mathcal{P}$ for in-clear, $\mathcal{A}$ for arithmetic circuits, and
$\mathcal{B}$ for boolean circuits). The
form $\widetilde{e} \rhd m$ denotes coercing the value of $\widetilde{e}$ to be
$m$-secret shared, which is useful for share-conversions between boolean and arithmetic and also for sharing public values. 
%The statements $\widetilde{s}$ are analogous to $s$.

@@

	\inferrule*[lab={\footnotesize{T-Read}}]
               {
                 \Gamma \vdash x : \sigma^{\ell}[n] \leadsto x\\\\
                 \Gamma \vdash e : \kw{uint}^{\mathcal{P}} \leadsto \widetilde{e}\\\\
                  \models e < n
               }
               {
                 \Gamma \vdash x[e] : \sigma^{\ell} \leadsto x[\widetilde{e}]
               }

@@

     \inferrule*[lab={\footnotesize{T-Write}}]
               {
                 \Gamma \vdash x : \sigma^{\ell}[n] \leadsto x\\\\
                 \Gamma \vdash e_{1} : \kw{uint}^{\mathcal{P}} \leadsto \widetilde{e_{1}}\\\\
                 \Gamma \vdash e_{2} : \sigma^{\ell} \leadsto \widetilde{e_{2}}\\\\
                  \models e_{1} < n
               }
               {
                 \Gamma \vdash x[e_{1}] := e_{2} \leadsto x[\widetilde{e_{1}}] := \widetilde{e_{2}} \mid \Gamma
               }

@@

\subsubsection*{Source to intermediate compilation judgments} We provide the compilation
rules in \figureref{compile}. We present the rules in a
declarative style, where the rules are non-syntax
directed, and the labels $\ell$ are chosen
non-deterministically. \sectionref{impl} describes the label  inference
scheme of our implementation.

The judgment $\Gamma \vdash e : \tau \leadsto \widetilde{e}$, where
$\Gamma$ maps variables $x$ to types $\tau$, says that
under $\Gamma$, $e$ (in source language) compiles to $\widetilde{e}$ (in the intermediate language) and $\widetilde{e}$  has type $\tau$.
The rule ({\sc{T-Cons}}) assigns the label $\mathcal{P}$ to the
constants, as the constants are always public. Rule ({\sc{T-Mult}})
compiles a multiplication to either a public multiplication  ($\mulop_{\mathcal{P}}$), or a
secret arithmetic multiplication ($\mulop_{\mathcal{A}}$). As
our compiler is cryptographic cost aware, it never compiles the
multiplication to boolean multiplication $\mulop_{\mathcal{B}}$. In a similar manner,
rule ({\sc{T-Gt}}) compiles $e_{1} > e_{2}$ to either public comparison,
or secret boolean comparison $>_\mathcal{B}$ (and never $>_{\mathcal{A}}$). The rule for conditional
({\sc{T-Cond}}) has two cases: when the conditional expression $e$ is of
type $\kw{bool}^{\mathcal{P}}$, both the branches have a base type
$\sigma^{\ell'}$, for an arbitrary $\ell'$, and the conditional is
compiled to a public conditional, whereas when the conditional
expression has type $\kw{bool}^{\mathcal{B}}$, $\ell'$ is also
$\mathcal{B}$, and the conditional is compiled to a secret conditional
using a boolean circuit. Note that we restrict the type of the branches
to be of base type. 
Rule ({\sc{T-Read}}) type checks an array read. It checks
that the array index $e$ is public, and uses a static bounds checking
judgment $ \models e < n$ to prove that the array index is in
bounds. Section~\ref{sec:impl} discusses our implementation of this check. Rule ({\sc{T-Inp}}) picks a label $m$ for the input. Finally,
the rule ({\sc{T-Sub}}) is the subsumption rule that coerces an expression of
type $\sigma^{\ell}$ to an expression of type $\sigma^{m}$ using the
coerce expression. It is important for security that the secrets cannot be coerced to public values and indeed ({\sc{T-Sub}}) does not permit it.
